\begin{appendices}
\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

%  \chapter{First appendix}
%  \section{First section}
%  \section{Second section}

  %\chapter{Second appendix}
  %\section{First section}
  %\section{Second section}
%\end{appendices}


\chapter{The deformation gradient}
\ref{appendix:defgrad}
%\section{The deformation gradient}
Deformation is a major property of interest when a continuum is influenced by external and internal forces.  The deformation results in relative change of position of material particles, called \textit{strain}. and is the primary property that causes and describe \textit{stress}. Strain is purely an observation, and it is not dependent on the material of interest. However one expects that a material undergoing strain, will give  forces within a continuum due to neighboring material particles interacting with one another. Therefore one derive material specific models to describe how a certain material will react to a certain amount of strain. These strain measures are used to define models for \textit{stress}, which is responsible for the deformation in materials \cite{Holzapfel2000}. Stress is defined as the internal forces that particles within a continuous material exert on each other, with dimension force per unit area.  \\

The equations of continuum mechanics can be derived with respect to either a deformed or undeformend configuration. The choice of refering our equations to the current or reference configuration is indifferent from a theoretical point of view. In practice however this choice can have a severe impact on our strategy of solution methods and physical of modelling.   \cite{Wriggers2006}. Regardless of configuration, the \textit{deformation gradient} and \textit{determinant of the deformation gradient} are essential measurement in structure mechanics. 
By \cite{Richter2016}, both configurations are considered.

\subsubsection*{Reference configuration}
\begin{defn}
Let $\hat{u}$ be a differential deformation field in the \textit{reference} configuration, $I$ be the Identity matrix and
the gradient $\hat{\nabla} = (\frac{\partial}{\partial x}, \frac{\partial}{\partial y}, \frac{\partial}{\partial z}) $. Then the \textit{deformation gradient} is given by,
\begin{align}
\bat{F} = I + \hat{\nabla} \bat{u} 
\end{align} 
\textit{expressing the local change of relative position under deformation.}
\end{defn}

\begin{defn}
Let $\hat{u}$ be a differential deformation field in the \textit{reference} configuration, $I$ be the Identity matrix and
the gradient $\hat{\nabla} = (\frac{\partial}{\partial x}, \frac{\partial}{\partial y}, \frac{\partial}{\partial z}) $. Then the \textit{determinant of the deformation gradient} is given by,
\begin{align}
J = det(\bat{F}) = det( I + \hat{\nabla} \bat{u} )
\end{align} 
\textit{expressing the local change of volume the configuration.}
\end{defn}

From the assumption of linear operator \textbf{F}, and no two particles $\bat{x}_a, \bat{x}_b \in \ha{V}$ occupy the same location for some time $V(t)$,  J to be greater than 0 \cite{Wriggers2006}. \\ \\

\subsubsection*{Current configuration}
\begin{defn}
Let $\mathbf{u}$ be a differential deformation field in the \textit{reference} configuration, $I$ be the Identity matrix and
the gradient $\mathbf{\nabla} = (\frac{\partial}{\partial x}, \frac{\partial}{\partial y}, \frac{\partial}{\partial z}) $. Then the \textit{deformation gradient} is given by,
\begin{align}
\mathbf{F} = I - \mathbf{\nabla} \mathbf{u} 
\end{align} 
\textit{expressing the local change of relative position under deformation.}
\end{defn}

\begin{defn}
Let $\mathbf{u}$ be a differential deformation field in the \textit{reference} configuration, $I$ be the Identity matrix and
the gradient $\mathbf{\nabla} = (\frac{\partial}{\partial x}, \frac{\partial}{\partial y}, \frac{\partial}{\partial z}) $. Then the \textit{determinant of the deformation gradient} is given by,
\begin{align}
J = det(\mathbf{F}) = det( I - \mathbf{\nabla} \mathbf{u} )
\end{align} 
\textit{expressing the local change of volume the configuration.}
\end{defn}


\chapter{Strain tensor}
\ref{appendix:strainapp}
The equations describing forces on our domain can be derived in accordinance with the current or reference configuration. With this in mind, different measures of strain can be derived with respect to which configuration we are interested in. We will here by \cite{Richter2016} show the most common measures of strain. We will first introduce the right \textit{Cauchy-Green} tensor \textbf{C}, which is one of the most used strain measures \cite{Wriggers2006}. \\ Uttrykk 1.3 fra Godboka, LAG TEGNING \\ 

Let $\ha{x}, \ha{y} \in \ha{V}$ be two points in our referemce configuration and let $\ha{a} = \ha{y} - \ha{x}$ denote the
length of the line bewtween these two points. As our domain undergoes deformation let 
$x = \ha{x} + \hat{u}( \ha{x} ) $ and $x = \ha{y} + \hat{u}( \ha{y} )  $ be the position of our points in the current configuration, and let $a = y - x$ be our new line segment. By \cite{Richter2016} we have by first order Taylor expansion

\begin{align*}
&y - x = \ha{y} + \hat{u}(\ha{y}) - \ha{x} - \hat{u}(\ha{x}) = \
\ha{y} - \ha{x} + \hat{\nabla}\ha{u}(\ha{x}) (\ha{y} - \ha{x}) 
+ \mathcal{O}(|\ha{y} - \ha{x} |^2) \\
&\frac{y - x}{|\ha{y} - \ha{x}|} = [I + \hat{\nabla}\hat{u}(\ha{x} ]  
\frac{\ha{y} - \ha{x}}{|\ha{y} - \ha{x}|} + \mathcal{O}(|\ha{y} - \ha{x} |) 
\end{align*}

This detour from \cite{Richter2016}  we have that 
\begin{align*}
&a = y - x = \hat{F}(\ha{x})\ha{a} +  \mathcal{O}(|\ha{a} |^2) \\
&|a| = \sqrt{ (\hat{F}\ha{a},\hat{F}\ha{a})+ \mathcal{O} (|\ha{a}^3|)  } = 
 \sqrt{ (\ha{a}^T, \hat{F}^T\hat{F}\ha{a})} + \mathcal{O} (|\ha{a}^2|)  
\end{align*}

We let $\ha{C} = \ha{F}^T \ha{F}$ denote the right \textit{Cauchy-Green tensor}.
By observation the Cauchy-Green tensor is not zero at the reference configuration 
\begin{align*}
\ha{C} =  \ha{F}^T \ha{F} = (I + \hat{\nabla} \ha{u})^T (I + \hat{\nabla} \ha{u}) = 1
\end{align*}

Hence it is convenient to introduce a tensor which is zero at the reference configuration. We define the \textit{Grenn-Lagrange strain tensor}, which arises from the squard rate of change of the linesegment \ha{a} and \textit{a}. By using the definition of the Cauchy-Green tensor we have the relation
\begin{align*}
&\frac{1}{2}(|a|^2 + |\ha{a}|^2) = \frac{1}{2}(\ha{a}^T\hat{C}\ha{a}
 -\ha{a}^T \ha{a} ) + \mathcal{O}(|\ha{a}^3| = 
 \ha{a}^T \big(\frac{1}{2} (\hat{F}^T \hat{F} - I) \big) \ha{a} 
 + \mathcal{O}(\ha{a}^3) \\
&\hat{E} = \frac{1}{2}(\hat{C} - I)
\end{align*}

Both the \textit{right Cauchy-Green tensor} $\hat{C}$ and the \textit{Green-Lagrange} $\hat{E}$ are refered to the Lagrangian coordinate system, hence the \textit{reference configuration}. \\
Using similar arguments (see \cite{Richter2016}, compsda) Eulerian counterparts of the Lagrangian stress tensors can be derived.

The \textit{left Cauchy-Green} strain tensor 
\begin{align*}
\mathbf{b} = \ha{F} \ha{F}^T = 
\end{align*}
and the \textit{Euler-Almansi} strain tensor
\begin{align*}
\mathbf{e} = \frac{1}{2} (I - \hat{F}^{-1}\hat{F}^{-T}) = \hat{F}^{-1}\hat{E}\hat{F}^{T}
\end{align*}

\end{appendices}